Magnetised Polish doughnuts revisited

### 2017-07-10 paper_revised_v2/paper_revised_v2.tex
removed boldface

The investigation of this type of systems, either by analytical or numerical means, may rely on the ability to construct suitable 
representations based on physical assumptions. The construction of equilibrium models of stationary discs around black holes has indeed a long tradition (see~\cite{Abramowicz:2013} and references therein). In particular, the so-called ``Polish doughnuts"~\citep{Abramowicz:1978,Kozlowski:1978} provide a  very general method to build equilibrium configurations of perfect fluid matter orbiting around a Kerr black hole. This fully relativistic model assumes that the disc is non-magnetised and that the matter obeys a  constant specific angular momentum distribution. This method was later extended by~\cite{Komissarov:2006}  by adding a purely azimuthal magnetic field to build magnetised tori around rotating black holes. {\bf Dynamical evolutions of magnetized tori built with the Komissarov solution were first reported by~\cite{Montero:2007}.}
On the other hand, assuming different distributions of angular momentum in the discs, \citet{Qian:2009} presented a method to build sequences of black hole thick accretion discs in dynamical equilibria, restricted however to the purely hydrodynamical case.

@@

\cite{Jaroszynski:1980} argued that the slope of the specific angular momentum should range between two limiting cases, namely $l = \mathrm{const.}$ and $\Omega = \mathrm{const}$. Following \citet{Qian:2009} we assume an angular momentum distribution ansatz given by  
\begin{equation}
l (r,\theta) = \left\{ \label{eq:ansatz} 
  \begin{array}{lr}
    l_0 \left(\frac{l_{\mathrm{K}}(r)}{l_0}\right)^{\beta}\sin^{2\gamma}{\theta} &  \text{for } r \geq r_{\mathrm{ms}}\\
    l_{\mathrm{ms}}(r)\sin^{2\gamma}{\theta} & \text{for } r < r_{\mathrm{ms}}
  \end{array}
\right.
\end{equation}
where constants $l_0$ and $l_{\mathrm{ms}}(r)$ are defined by $l_0 \equiv \eta l_{\mathrm{K}}(r_{\mathrm{ms}})$ and $l_{\mathrm{ms}}(r) \equiv l_0 [l_{\mathrm{K}}(r_{\mathrm{ms}})/l_0]^{\beta}$. Therefore, the model for the distribution of angular momentum has three free parameters, $\beta$ , $\gamma$ and $\eta$, whose range of variation is given by~\citep{Qian:2009}
\begin{equation}
0 \leq \beta \leq 1, \quad -1 \leq \gamma \leq 1, \quad 1 \leq \eta \leq \eta_{\mathrm{max}},
\end{equation}
with $\eta_{\mathrm{max}} = l_{\mathrm{K}}(r_{\mathrm{mb}})/l_{\mathrm{K}}(r_{\mathrm{ms}})$. In this paper, and as it is done for hydrodynamical discs in~\citet{Qian:2009}, we choose $\eta = \eta_{\mathrm{max}}$, and then we can write $l_0$ as $l_0 = l_{\mathrm{K}}(r_{\mathrm{mb}})$. For this choice of $\eta$, we can find the location of the cusp of the disc  within the range $r_{\mathrm{mb}} \leq r_{\mathrm{cusp}} \leq r_{\mathrm{ms}}$ (for $0 \leq \beta \leq 1$). {\bf The cusp is defined as the circle in the equatorial plane on which the pressure gradient vanishes and the angular momentum of the disc equals the Keplerian angular momentum.} Moreover, this value of $\eta$ guarantees that constant angular momentum discs ($\beta = \gamma = 0$) with their inner edge located at the cusp ($r_{\mathrm{in}} = r_{\mathrm{cusp}}$) have no outer boundary, i.e~they are infinite discs (see also~\cite{Font:2002}).

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the {\bf fluid enthalpy density}and {\bf fluid} pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogue equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

@@

To construct our models of magnetised discs we follow the approach described in \citet{Qian:2009}. First, we find the radial location of the cusp and of the centre of the disc in the equatorial plane, $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$, defined as the solutions to the equation $l(r) - l_{\mathrm{K}} = 0$.
Next, we compute the partial derivatives of the potential, Eq.~\eqref{eq:potential}
\begin{equation}\label{eq:radial_der_pot}
\partial_r W = \partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\,,
\end{equation}
and
\begin{equation}\label{eq:polar_der_pot}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\,.
\end{equation}
Then, we integrate the radial partial derivative of the potential along the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ (assuming $W_{\mathrm{cusp}} = 0$) at the equatorial plane, thus obtaining the equatorial distribution of the potential between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$
\begin{equation}\label{eq:equatorial_pot}
W_{\mathrm{eq}}(r) = \int^{r_{\mathrm{c}}}_{r_{\mathrm{cusp}}}\left(\partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\right).
\end{equation}
Following~\citet{Qian:2009} (see also~\citet{Jaroszynski:1980}) we can divide equations \eqref{eq:radial_der_pot} and \eqref{eq:polar_der_pot} to obtain
\begin{equation}\label{eq:F}
F(r, \theta) = -\frac{\partial_r W}{\partial_{\theta} W} = \frac{\mathrm{d}\theta}{\mathrm{d}r}\,,
\end{equation}
where function $F(r, \theta)$ is known in closed form for the Kerr metric once an assumption for the angular momentum distribution has been made (cf.~Eq.~(\ref{eq:ansatz})). For this reason, Eq.~(\ref{eq:F}) also takes the form of an ordinary differential equation for the surfaces of constant potential, $\theta=\theta(r)$, which, upon integration, yields the location of those surfaces. Note that in~\citet{Qian:2009} these are surfaces of constant {\it {\bf fluid} pressure} instead, since their discs are purely hydrodynamical.

\begin{figure*}[t]
\centering
\includegraphics[scale=0.14]{figures/fig1a.pdf}
\hspace{-0.3cm}
\includegraphics[scale=0.14]{figures/fig1b.pdf}
\hspace{-0.2cm}
\includegraphics[scale=0.14]{figures/fig1c.pdf}
\caption{Comparison with Komissarov's solution for a constant angular momentum model. The left panel shows the {\bf rest-mass} density distribution in logarithm scale for all radii and all angles while the middle and right panels show, respectively, the angular profile of the logarithm of the {\bf rest-mass} density at the centre of the disc and the radial profile of the {\bf rest-mass} density at the equatorial plane.}
\label{komissarov}
\end{figure*}

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the {\bf fluid} pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the {\bf enthalpy density} at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the {\bf enthalpy density} distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the rest-mass density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. We note that the fluid {\bf enthalpy density} $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use poytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implictly assuming that $h=1$, i.e.~the discs we build in our procedure are non-relativistic from a thermodynamical point of view.
%{\bf We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
%this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
%$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$.}

For the integration of Eq.~\eqref{eq:equatorial_pot} we use the composite Simpson's rule. It is important to use a very small integration step because the slope is very steep. In this work, we use a step $\Delta r = 10^{-6}$. Using the analytic, constant angular momentum case for comparison, we tested that 
a larger value of $\Delta r$ gives unacceptable accuracy losses. On the other hand, to integrate the ordinary differential equation \eqref{eq:F} we use a fourth-order Runge-Kutta method. Again, it is also important here to choose a suitable step of integration, especially at the outer end of the disc, because Eq.~\eqref{eq:F} diverges at the equatorial plane (the equipotential surfaces cross the equatorial plane perpendicularly). {\bf We show the proof of this statment in Appendix~\ref{div_partial_W}.}

\begin{table}
\caption{List of models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$. From left to right the columns report: the model name (with A, B and C standing for black hole spin $a = 0.5$, $0.9$ and $0.99$, respectively), the parameters of the angular momentum distribution $\beta$ and $\gamma$, the corresponding radii of the maximum {\bf fluid} pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, the gravitational potential difference between the centre and the inner radius, $\Delta W\equiv W_{\rm c}-W_{\rm in}$, and the inner and outer radii of the discs, $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$.}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $7.15$ &  $8.14$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $7.15$ &  $7.65$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $7.15$ &  $7.45$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $7.15$ &  $8.32$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $7.15$ &  $7.28$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.59$ &  $3.98$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.59$ &  $3.83$  & $-4.32$ & $1.86$ & $5.35$\\ 
B3 & $0.9$ & $0.9$ & $3.59$ &  $3.73$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.59$ &  $3.85$  & $-12.9$ & $1.73$ & $4.54$\\ 
B5 & $0.9$ & $0.0$ & $3.59$ &  $3.81$  & $-0.54$ & $2.08$ & -- \\  
 \hline 
C1 & $0.0$ & $0.0$ & $1.98$ &  $2.28$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.98$ &  $2.09$  & $-7.34$ & $1.26$ & $2.50$\\ 
C3 & $0.9$ & $0.9$ & $1.98$ &  $2.04$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.98$ &  $2.10$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.98$ &  $2.08$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}

>>>>

The investigation of this type of systems, either by analytical or numerical means, may rely on the ability to construct suitable 
representations based on physical assumptions. The construction of equilibrium models of stationary discs around black holes has indeed a long tradition (see~\cite{Abramowicz:2013} and references therein). In particular, the so-called ``Polish doughnuts"~\citep{Abramowicz:1978,Kozlowski:1978} provide a  very general method to build equilibrium configurations of perfect fluid matter orbiting around a Kerr black hole. This fully relativistic model assumes that the disc is non-magnetised and that the matter obeys a  constant specific angular momentum distribution. This method was later extended by~\cite{Komissarov:2006}  by adding a purely azimuthal magnetic field to build magnetised tori around rotating black holes. Dynamical evolutions of magnetized tori built with the Komissarov solution were first reported by~\cite{Montero:2007}.
On the other hand, assuming different distributions of angular momentum in the discs, \citet{Qian:2009} presented a method to build sequences of black hole thick accretion discs in dynamical equilibria, restricted however to the purely hydrodynamical case.

@@

\cite{Jaroszynski:1980} argued that the slope of the specific angular momentum should range between two limiting cases, namely $l = \mathrm{const.}$ and $\Omega = \mathrm{const}$. Following \citet{Qian:2009} we assume an angular momentum distribution ansatz given by  
\begin{equation}
l (r,\theta) = \left\{ \label{eq:ansatz} 
  \begin{array}{lr}
    l_0 \left(\frac{l_{\mathrm{K}}(r)}{l_0}\right)^{\beta}\sin^{2\gamma}{\theta} &  \text{for } r \geq r_{\mathrm{ms}}\\
    l_{\mathrm{ms}}(r)\sin^{2\gamma}{\theta} & \text{for } r < r_{\mathrm{ms}}
  \end{array}
\right.
\end{equation}
where constants $l_0$ and $l_{\mathrm{ms}}(r)$ are defined by $l_0 \equiv \eta l_{\mathrm{K}}(r_{\mathrm{ms}})$ and $l_{\mathrm{ms}}(r) \equiv l_0 [l_{\mathrm{K}}(r_{\mathrm{ms}})/l_0]^{\beta}$. Therefore, the model for the distribution of angular momentum has three free parameters, $\beta$ , $\gamma$ and $\eta$, whose range of variation is given by~\citep{Qian:2009}
\begin{equation}
0 \leq \beta \leq 1, \quad -1 \leq \gamma \leq 1, \quad 1 \leq \eta \leq \eta_{\mathrm{max}},
\end{equation}
with $\eta_{\mathrm{max}} = l_{\mathrm{K}}(r_{\mathrm{mb}})/l_{\mathrm{K}}(r_{\mathrm{ms}})$. In this paper, and as it is done for hydrodynamical discs in~\citet{Qian:2009}, we choose $\eta = \eta_{\mathrm{max}}$, and then we can write $l_0$ as $l_0 = l_{\mathrm{K}}(r_{\mathrm{mb}})$. For this choice of $\eta$, we can find the location of the cusp of the disc  within the range $r_{\mathrm{mb}} \leq r_{\mathrm{cusp}} \leq r_{\mathrm{ms}}$ (for $0 \leq \beta \leq 1$). The cusp is defined as the circle in the equatorial plane on which the pressure gradient vanishes and the angular momentum of the disc equals the Keplerian angular momentum. Moreover, this value of $\eta$ guarantees that constant angular momentum discs ($\beta = \gamma = 0$) with their inner edge located at the cusp ($r_{\mathrm{in}} = r_{\mathrm{cusp}}$) have no outer boundary, i.e~they are infinite discs (see also~\cite{Font:2002}).

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the {\bf fluid enthalpy density} and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogue equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

@@

To construct our models of magnetised discs we follow the approach described in \citet{Qian:2009}. First, we find the radial location of the cusp and of the centre of the disc in the equatorial plane, $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$, defined as the solutions to the equation $l(r) - l_{\mathrm{K}} = 0$.
Next, we compute the partial derivatives of the potential, Eq.~\eqref{eq:potential}
\begin{equation}\label{eq:radial_der_pot}
\partial_r W = \partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\,,
\end{equation}
and
\begin{equation}\label{eq:polar_der_pot}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\,.
\end{equation}
Then, we integrate the radial partial derivative of the potential along the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ (assuming $W_{\mathrm{cusp}} = 0$) at the equatorial plane, thus obtaining the equatorial distribution of the potential between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$
\begin{equation}\label{eq:equatorial_pot}
W_{\mathrm{eq}}(r) = \int^{r_{\mathrm{c}}}_{r_{\mathrm{cusp}}}\left(\partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\right).
\end{equation}
Following~\citet{Qian:2009} (see also~\citet{Jaroszynski:1980}) we can divide equations \eqref{eq:radial_der_pot} and \eqref{eq:polar_der_pot} to obtain
\begin{equation}\label{eq:F}
F(r, \theta) = -\frac{\partial_r W}{\partial_{\theta} W} = \frac{\mathrm{d}\theta}{\mathrm{d}r}\,,
\end{equation}
where function $F(r, \theta)$ is known in closed form for the Kerr metric once an assumption for the angular momentum distribution has been made (cf.~Eq.~(\ref{eq:ansatz})). For this reason, Eq.~(\ref{eq:F}) also takes the form of an ordinary differential equation for the surfaces of constant potential, $\theta=\theta(r)$, which, upon integration, yields the location of those surfaces. Note that in~\citet{Qian:2009} these are surfaces of constant {\it fluid pressure} instead, since their discs are purely hydrodynamical.

\begin{figure*}[t]
\centering
\includegraphics[scale=0.14]{figures/fig1a.pdf}
\hspace{-0.3cm}
\includegraphics[scale=0.14]{figures/fig1b.pdf}
\hspace{-0.2cm}
\includegraphics[scale=0.14]{figures/fig1c.pdf}
\caption{Comparison with Komissarov's solution for a constant angular momentum model. The left panel shows the rest-mass density distribution in logarithm scale for all radii and all angles while the middle and right panels show, respectively, the angular profile of the logarithm of the rest-mass density at the centre of the disc and the radial profile of the rest-mass density at the equatorial plane.}
\label{komissarov}
\end{figure*}

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the fluid pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the {\bf enthalpy density} at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the {\bf enthalpy density} distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the rest-mass density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. We note that the fluid {\bf enthalpy density} $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use poytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implictly assuming that $h=1$, i.e.~the discs we build in our procedure are non-relativistic from a thermodynamical point of view.
%{\bf We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
%this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
%$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$.}

For the integration of Eq.~\eqref{eq:equatorial_pot} we use the composite Simpson's rule. It is important to use a very small integration step because the slope is very steep. In this work, we use a step $\Delta r = 10^{-6}$. Using the analytic, constant angular momentum case for comparison, we tested that 
a larger value of $\Delta r$ gives unacceptable accuracy losses. On the other hand, to integrate the ordinary differential equation \eqref{eq:F} we use a fourth-order Runge-Kutta method. Again, it is also important here to choose a suitable step of integration, especially at the outer end of the disc, because Eq.~\eqref{eq:F} diverges at the equatorial plane (the equipotential surfaces cross the equatorial plane perpendicularly). We show the proof of this statment in Appendix~\ref{div_partial_W}.

\begin{table}
\caption{List of models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$. From left to right the columns report: the model name (with A, B and C standing for black hole spin $a = 0.5$, $0.9$ and $0.99$, respectively), the parameters of the angular momentum distribution $\beta$ and $\gamma$, the corresponding radii of the maximum fluid pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, the gravitational potential difference between the centre and the inner radius, $\Delta W\equiv W_{\rm c}-W_{\rm in}$, and the inner and outer radii of the discs, $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$.}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $7.15$ &  $8.14$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $7.15$ &  $7.65$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $7.15$ &  $7.45$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $7.15$ &  $8.32$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $7.15$ &  $7.28$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.59$ &  $3.98$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.59$ &  $3.83$  & $-4.32$ & $1.86$ & $5.35$\\ 
B3 & $0.9$ & $0.9$ & $3.59$ &  $3.73$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.59$ &  $3.85$  & $-12.9$ & $1.73$ & $4.54$\\ 
B5 & $0.9$ & $0.0$ & $3.59$ &  $3.81$  & $-0.54$ & $2.08$ & -- \\  
 \hline 
C1 & $0.0$ & $0.0$ & $1.98$ &  $2.28$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.98$ &  $2.09$  & $-7.34$ & $1.26$ & $2.50$\\ 
C3 & $0.9$ & $0.9$ & $1.98$ &  $2.04$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.98$ &  $2.10$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.98$ &  $2.08$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}
